---
title: "Testing the role of the generative cause in prevenitve relations"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r, message = FALSE, echo = FALSE, warning=FALSE}
# packages
library(ez)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(overlapping) # to compute the 
library(tidyr)
library(showtext)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```



# Subject Demographics

```{r}
# read data 
tdata <- read.delim("Exp2_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)




# Demographics
#demogr_data <- tdata[!duplicated(tdata$sID), ]

mean(tdata$Age)
sd(tdata$Age)
min(tdata$Age)
max(tdata$Age)
table(tdata$Sex)


count(tdata, c("danger", "order"))
```
# Data preparation

```{r}
# declare factors and their levels 
#tdata$Effect_info <- factor(tdata$Effect_info, levels = c("full_info", "only_effect"), labels = c("full_info", "only_effect"))

tdata$danger <- factor(tdata$danger, levels = c("weakest", "weak", "strong"), labels = c("qa = 0.25", "qa = 0.50", "qa = 1.0"))

tdata$order <- factor(tdata$order, levels = c("non_attack_first","attack_first"), labels = c("non_attack_first","attacke_first"))

#subset for types of ratings
singular_ratings <- subset(tdata, select = c(1:3,singular_query_a, singular_query_na))


# subset for strength ratings 
strength_ratings <- subset(tdata, select = c(1:3,preventive_strength, danger_strength))


# turning into long format
singular_ratings <- singular_ratings %>% gather(Test_case, Rating, singular_query_a:singular_query_na, factor_key=TRUE)

strength_ratings <- strength_ratings %>% gather(Variable, Rating, preventive_strength:danger_strength, factor_key=TRUE)



# transform ratings 
singular_ratings$Rating <- (singular_ratings$Rating - 1) * 0.10
strength_ratings$Rating <- (strength_ratings$Rating - 1) * 0.10


strength_ratings$Variable <- factor(strength_ratings$Variable, levels = c("preventive_strength","danger_strength"), labels = c("Preventive cause","Generative cause"))

singular_ratings$Test_case <- factor(singular_ratings$Test_case, levels = c("singular_query_a","singular_query_na"), labels = c("attacked","non-attacked"))


```


# Graphs

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(face = "bold", size = 20),
        axis.title.y = element_text(face = "bold", size = 20),
        axis.text.x = element_text(size = 18, angle = 0), 
        axis.text.y = element_text(size = 14, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


library(see)
## first, turn sID into a factor
strength_ratings$sID <- factor(strength_ratings$sID)

pd <- position_dodge(width = 0.3)

strength_ratings$valueJitter <- jitter(strength_ratings$Rating, factor = 1, amount = 0.04)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 
#effects.labs <- c("Preventive cause", "Generative cause")
#names(effects.labs) <- c("same domain", "different domains")

palette <- c("#FFAE49", "#8cb495", "#12658b")


g <- ggplot(strength_ratings, aes(x=danger, y=valueJitter, group=sID)) +
  #guides(fill=FALSE)+
  #facet_grid( ~ type_effect, labeller = labeller(type_effect =effects.labs))+
  facet_grid( ~ Variable)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-0.05, 1.05), breaks=seq(0, 1.00, 0.10), expand = c(0,0)) +
  scale_x_discrete(labels=c("qa = 0.25", "qa = 0.5", "qa = 1.0")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", alpha =0.5) +
  geom_violinhalf(aes(y = Rating, group = danger), color = NA, fill = "#F8766D", position=position_dodge(1), alpha = 0.5)+
  #geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_point(position = pd, alpha = 0.2, color = "#F8766D", alpha = 0.6) +
  #geom_jitter()+
  stat_summary(aes(y = Rating, group = 1), fun.data = mean_cl_boot, geom = "errorbar", width = 0, size = 1) +
  #stat_summary(aes(y = Rating,group=1), fun.y=mean, colour="black", geom="line",group=1, size = 1.5, linetype = "solid", alpha = 1)+
  stat_summary(aes(y = Rating), fun.y=mean, geom="point", color = "black", fill = "#F8766D", shape = 22, size = 5, group=1, alpha = 1)+
  stat_summary(aes(y = Rating), fun.y=median, geom="point", color = "black", shape = 3, size = 4, group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Strength of generative cause", y = "Causal Strength Rating") +
  scale_color_manual(name = "Entity",values=palette)+
  scale_fill_manual(name = "Entity",values=palette)+
  theme(legend.position = "none")+
  myTheme
g


#ggsave("results_strengths.svg",width=10,height=5)
#ggsave("results_strengths.pdf",width=10,height=5)
```


```{r}
library(tidyverse)
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(face = "bold", size = 20),
        axis.title.y = element_text(face = "bold", size = 20),
        axis.text.x = element_text(size = 15, angle = 0), 
        axis.text.y = element_text(size = 14, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


library(see)
## first, turn sID into a factor
singular_ratings$sID <- factor(strength_ratings$sID)

pd <- position_dodge(width = 0.3)

singular_ratings$valueJitter <- jitter(strength_ratings$Rating, factor = 1, amount = 0.04)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

# new labes for the facets 
#effects.labs <- c("Preventive cause", "Generative cause")
#names(effects.labs) <- c("same domain", "different domains")

palette <- c("#FFAE49", "#8cb495", "#12658b")


g <- ggplot(singular_ratings, aes(x=Test_case, y=valueJitter, group=sID)) +
  #guides(fill=FALSE)+
  #facet_grid( ~ type_effect, labeller = labeller(type_effect =effects.labs))+
  facet_grid( ~ danger)+
  #ggtitle("Subjects' causal srength ratings") +
  scale_y_continuous(limits = c(-0.05, 1.05), breaks=seq(0, 1.00, 0.10), expand = c(0,0)) +
  scale_x_discrete(labels=c("[c+, e-, a+]", "[c+, e-, a-]")) +
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 0.25"),
             aes(yintercept = 0.721), col = "black", linetype = "dashed")+
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 0.50"),
             aes(yintercept = 0.777), col = "black", linetype = "dashed")+
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 1.0"),
             aes(yintercept = 0.881), col = "black", linetype = "dashed")+
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 0.25"),
             aes(yintercept = 0.452), col = "black", linetype = "dashed")+
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 0.50"),
             aes(yintercept = 0.428), col = "black", linetype = "dashed")+
  geom_hline(data = singular_ratings %>% filter(danger == "qa = 1.0"),
             aes(yintercept = 0.445), col = "black", linetype = "dashed")+
  geom_violinhalf(aes(y = Rating, group = Test_case), fill = "#619CFF", color = NA, position=position_dodge(1), alpha = 0.4)+
  geom_line(position = pd, color = "black", size = 1, alpha=0.04) +
  geom_point(color = "#619CFF", position = pd, alpha = 0.2) +
  stat_summary(aes(y = Rating, group = 1), fun.data = mean_cl_boot, geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = Rating,group=1), fun.y=mean, colour="#619CFF", geom="line",group=1, size = 1.5, linetype = "solid", alpha = 1)+
  stat_summary(aes(y = Rating), fill = "#619CFF", fun.y=mean, geom="point", color = "black", shape = 22, size = 4, group=1, alpha = 1)+
  stat_summary(aes(y = Rating), fun.y = median, geom="point", color = "black", shape = 3, size = 4, group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Singular Test Case", y = "Singular Prevention Rating") +
  scale_color_manual(name = "Entity",values=palette)+
  scale_fill_manual(name = "Entity",values=palette)+
  theme(legend.position = "none")+
  myTheme
g


#ggsave("results_lines_singular.svg",width=10,height=5)
#ggsave("results_lines_singular.pdf",width=10,height=5)
```



# Statistical Analyses

## Strength ratings for preventer and generator


```{r, echo=FALSE}
library(afex)
a1 <- aov_car(Rating ~ danger * Variable + Error(sID/Variable), anova_table = list(es = "pes"), strength_ratings)
 a1
```



```{r}
# follow-up analysis to get means etc.
library(emmeans)
ls1 <- lsmeans(a1, c("danger", "Variable")) # joint evaluation (basically gives the same table)
ls1


# contrasts 
contrasts <- emmeans(a1, ~ danger*Variable)
s <- pairs(contrasts, adjust = "none")
s
```
No differences for the preventive strength queries. For the generative strength queries, ratings did not differ between qa = 0.25 and qa = 0.50, but ratings were significantly higher in the qa = 1.0 condition.



```{r}
# get descriptives tha include medians
library(vtable)
sum_data <- strength_ratings
sum_data$cond <- paste(sum_data$danger, sum_data$Variable)

sum <- st(sum_data, vars = c('Rating'), group = c("cond"), add.median=T, out='return', group.long=T)
sum

# see: https://cran.r-project.org/web/packages/vtable/vignettes/sumtable.html
```



## Singular Prevention ratings 

```{r, echo=FALSE}
library(afex)
a1 <- aov_car(Rating ~ danger * Test_case *order + Error(sID/Test_case), anova_table = list(es = "pes"), singular_ratings)
a1
```



```{r}
# follow-up analysis to get means etc.
library(emmeans)
ls1 <- lsmeans(a1, c("danger", "Test_case")) # joint evaluation (basically gives the same table)
ls1


# contrasts 
contrasts <- emmeans(a1, ~ danger*Test_case)
s <- pairs(contrasts, adjust = "none")
s
```

```{r}
# get descriptives tha include medians
library(vtable)
sum_data <- singular_ratings
sum_data$cond <- paste(sum_data$danger, sum_data$Test_case)

sum <- st(sum_data, vars = c('Rating'), group = c("cond"), add.median=T, out='return', group.long=T)
sum

# see: https://cran.r-project.org/web/packages/vtable/vignettes/sumtable.html
```

# Correlation analysis

An interesting question is whether subjects who gave very high qa ratings tended to give the highest singular prevention ratings. 


```{r}
cor.test(tdata$danger_strength, tdata$singular_query_a)
```






















