---
title: "Preemption in Prevention Suppl."
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---




```{r, message = FALSE, echo = FALSE, warning=FALSE}
# packages
library(ez)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(overlapping) # to compute the 
library(tidyr)
library(showtext)
#install.packages("expss")
library(expss)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```



# Subject Demographics

```{r}
# read data 
tdata <- read.delim("Exp_supp1_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)


# Demographics
#demogr_data <- tdata[!duplicated(tdata$sID), ]

mean(tdata$age)
sd(tdata$age)
min(tdata$age)
max(tdata$age)
table(tdata$gender)

```



# Data preparation

```{r}
# declare factors and their levels 
tdata$condition_1 <- factor(tdata$condition_1, levels = c(1, 2), labels = c("left", "right"))


# recode responses so that 1 = first, 2 = second, 3 = both
tdata$selection_rec <- c(1:length(tdata$run_id))
tdata$selection_rec[tdata$condition_1 =="left" & tdata$main_dv == 1] <- 1
tdata$selection_rec[tdata$condition_1 =="left" & tdata$main_dv == 2] <- 2
tdata$selection_rec[tdata$condition_1 =="left" & tdata$main_dv == 3] <- 3

tdata$selection_rec[tdata$condition_1 =="right" & tdata$main_dv == 1] <- 2
tdata$selection_rec[tdata$condition_1 =="right" & tdata$main_dv == 2] <- 1
tdata$selection_rec[tdata$condition_1 =="right" & tdata$main_dv == 3] <- 3




tdata$selection_verbal <- c(1:length(tdata$run_id))
tdata$selection_verbal[tdata$selection_rec == 1] <- "first"
tdata$selection_verbal[tdata$selection_rec == 2] <- "second"
tdata$selection_verbal[tdata$selection_rec == 3] <- "both"



```


## Get percentages 

What preventer is selected in the different prevention cases?

```{r}
# create a summary dataset that also contains the percentages
library(dplyr)
plotdata <- tdata %>%
  group_by(selection_verbal) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))

plotdata
```


## Conduct Proportion tests 

They yield the 95% CIs for the proportion estimates.


```{r}
# append CI of the proportions 
cases <- plotdata$n
N <- sum(plotdata$n)

t <- prop.test(cases[1],N, p = 1/3, correct=FALSE) # get CI for Invariance government
CI_both_min <- t$conf.int[1]
CI_both_max <- t$conf.int[2]
t


t <- prop.test(cases[2],N, p = 1/3, correct=FALSE) # get CI for Invariance government
CI_first_min <- t$conf.int[1]
CI_first_max <- t$conf.int[2]
t




CI_low <- c(CI_both_min, CI_first_min)

CI_up <- c(CI_both_max, CI_first_max)

plotdata$CI_low <- CI_low
plotdata$CI_up <- CI_up

```

Subjects could choose between three options (first, second, both). The majority chose the first preventer, with p = 0.6875. This probability is difference from chance (p = 0.33). 


# Graph of the Results 

Subjects' preventer selections.


```{r}
library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16, face = "bold"),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 16),
        legend.title = element_text(face = "bold", size = 16),
        strip.text.x = element_text(size = 16),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"),
        axis.ticks.x = element_blank())


g<- ggplot(plotdata, 
       aes(x = 0, y = pct,
           fill = factor(selection_verbal, 
                         levels = c("first", "second", "both"),
                         labels = c("first", "second", "both")))) +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(ylim = c(0, 1))+
  geom_bar(stat = "identity",
           position = "dodge") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent) +
  geom_text(aes(label = lbl), 
            size = 5, 
            position = position_dodge(width = .9),
            vjust = -5.7) +
  geom_errorbar(aes(ymin=CI_low, ymax=CI_up), width=.0,
                 position=position_dodge(.9)) +
  #geom_hline(yintercept=1/3, linetype="dashed", color = "red")+
  scale_fill_brewer(palette = "Set2") +
  labs(y = "Percent", 
       fill = "Singular \nPreventer",
       x = "Singular occasion")+
  myTheme

g
ggsave("Exp_suppl1_selections.svg",width=4,height=4)
```


